---
title: "Untitled"
author: "Jeff Suitor"
date: "11/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import libraries
```{r message=FALSE}
library("readr")
library("MASS")
library("car")
library("pscl")
library("caret")
```

Clean the data
```{r}
data_og = read_csv("heart.csv", show_col_types = FALSE)
data = data_og
data = subset(data, Oldpeak>=0) # Oldpeak shouldn't be negative
data = subset(data, RestingBP>0) # Resting BP can't be 0
data = subset(data, Cholesterol>0) # Cholesterol can't be 0

col_names = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina", "FastingBS")

data[,col_names] <- lapply(data[,col_names] , factor)

rm(list=c("col_names"))
```

Check for NA
```{r}
sum(is.na(data))
```

Create the model
```{r}
model = glm(HeartDisease ~ ., data = data, family = "binomial"(link="logit"))
```

There is negative impacts on on the different types of pain because ASY is baseline and other types of pain offer a protective effect.


Check for multiple logistic regression assumptions:

1. Appropriate Outcome Type -> Heart disease is either true or false so yes it is appropriate.
```{r}
length(unique(data$HeartDisease))
```

2. There is a linear relationship between the explanatory variable and the logit of the response variable
```{r}
logodds = model$linear.predictors
old_peak = (data$Oldpeak + 0.0000000000000001)
boxTidwell(logodds ~ data$Age)
boxTidwell(logodds ~ data$RestingBP)
boxTidwell(logodds ~ data$Cholesterol)
boxTidwell(logodds ~ data$MaxHR)
boxTidwell(logodds ~ old_peak)

plot(data$Age, logodds)
plot(data$RestingBP, logodds)
plot(data$Cholesterol, logodds)
plot(data$MaxHR, logodds, main="Max Heart Rate vs Log Odds", xlab="Max Heart Rate", ylab="Log Odds")
plot(data$Oldpeak, logodds,  main="Oldpeak vs Log Odds", xlab="Oldpeak", ylab="Log Odds")

rm(list=c("logodds", "old_peak"))
```

Recreate the model without the additional columns
```{r}
data = subset(data, select= -c(Oldpeak, MaxHR))
model2 = glm(HeartDisease ~ ., data = data, family = "binomial"(link="logit"))
```

3. There is no multicollinearity among explanatory variables -> All VIF values are less than 2 and many are close to 1 indicating that there is only moderate correlation with other predictor variables but that it is not severe enough to impact our model.
```{r}
vif(model2)
```

4. There are no extreme outliers
```{r}
cooks_distance <- cooks.distance(model2)

n = nrow(data)
plot(cooks_distance, main = "Cooks Distance for Influential Observations", pch=20)
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line
#identify influential points
influential_obs = as.numeric(names(cooks_distance)[(cooks_distance > (4/n))])

#define new data frame with influential points removed
data = data[-influential_obs, ]

rm(list=c("cooks_distance", "influential_obs", "n"))
```
Redo the model without the outliers
```{r}
model3 = glm(HeartDisease ~ ., data = data, family = "binomial"(link="logit"))
summary(model3)
```

Viewing the plots we can see there is no distinct pattern between the log odds and the params.

Run an anova using chi squared
```{r}
anova(model3, test="LRT")
```

Plot the model to see the residuals
```{r}
plot(model3)
summary(model3)
```

Run a sensitivity analysis
```{r}
predictions = ifelse(predict(model3, type="response") >= 0.5, 1, 0)
actual_values = model3$y
conf_matrix = table(predictions, actual_values)
conf_matrix
sensitivity(conf_matrix)
specificity(conf_matrix)
```

Run stepAIC to reduce the model
```{r}
model4 = stepAIC(model3)
```

Analyze the results
```{r}
summary(model4)
summary(model4$fitted.values)
```

Perform sensitivity and specificity
```{r}
predictions = ifelse(predict(model4, type="response") >= 0.5, 1, 0)
actual_values = model4$y
conf_matrix = table(predictions, actual_values)
conf_matrix
sensitivity(conf_matrix)
specificity(conf_matrix)
```

ANOVA comparison
```{r}
anova(model4, model3, test="LRT")
```



```{r}
mat = matrix(c(335, 52, 55, 304), ncol=2, byrow=TRUE)
conf_matrix = as.table(mat)
conf_matrix
sensitivity(conf_matrix) * 100
specificity(conf_matrix)* 100
```

